Poisson approximate likelihood compared to the particle filter

### 2024-08-30 ms_01.tex
output_01

\documentclass[10pt]{article}
\usepackage{color}
\usepackage{graphicx} % Required for inserting
\usepackage{booktabs} % For better-quality horizontal lines
\usepackage{multirow}
\usepackage{makecell}
% \usepackage{threeparttable} % For table notes
\usepackage{array} % For table formatting
\usepackage{subcaption}
\usepackage{setspace} % For double-line spacing
\usepackage{float} 
\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage[font={small,it}]{caption}
\usepackage[utf8]{inputenc}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{verbatim}
\usepackage{amsfonts,amssymb,amsthm}
\usepackage{url}
\usepackage[colorlinks=true,citecolor=blue]{hyperref}
% Using \doublespacing in the preamble 
% changes the text to double-line spacing
% \doublespacing

@@

\bibliographystyle{apalike}

\begin{document}

@@

<<setup,echo=FALSE,message=FALSE,warning=FALSE>>=
rm(list = ls())     # clear objects
library("knitr")
opts_knit$set(concordance=TRUE)
opts_chunk$set(
  tidy = TRUE,
  cache = TRUE,
#  cache = FALSE,  # set cache = FALSE when editing the code
# message = FALSE,
# warning = FALSE,
  tidy.opts = list(
    keep.blank.line = FALSE
  ),
  comment = "",
  echo = FALSE,
  results = FALSE,
  dev.args = list(
    bg = "transparent",
    pointsize = 9
  ),
  fig.path = "figure/"
)

myround <- function(x, digits = 1) {
  # taken from the broman package
  if (digits < 1)
    stop("This is intended for the case digits >= 1.")
  if (length(digits) > 1) {
    digits <- digits[1]
    warning("Using only digits[1]")
  }
  tmp <- sprintf(paste("%.", digits, "f", sep = ""), x)
  zero <- paste0("0.", paste(rep("0", digits), collapse = ""))
  tmp[tmp == paste0("-", zero)] <- zero
  tmp
}

graphics.off()      # close graphics windows
library(pomp)
library(ggplot2)
library(scales)
library(foreach)
library(doParallel)
library(doFuture)
library(dplyr)
library(Rcpp)
library(RcppArmadillo)
sourceCpp('rotavirus_normq.cpp')
source("choose_n_fits.R")
source("eval_logLik_pomp.R")

RUN_LEVEL = 1
NP_MIF       = switch(RUN_LEVEL, 100,   5000,  50000)
NMIF         = switch(RUN_LEVEL, 5,    20,    100)
ncores       = switch(RUN_LEVEL, 4,    10,    36)
top_n_fits   = switch(RUN_LEVEL, 1,    2,     12)
sim_rep      = switch(RUN_LEVEL, 5,    20,    100) #sim_ovov

#population number
pop <- 82372825
time <- c(1:416)

out_dir <- paste0("output_0",RUN_LEVEL, "/")
@

@@

<<OvOv-model, echo=F, warning=F, message=F>>=
# measurement model 
dmeas <- Csnippet("
  lik =  dbinom(cases1, H1, q1, 1) +
  dbinom(cases2, H2, q2, 1) +
  dbinom(cases3, H3, q3, 1);"
)
rmeas <-  Csnippet("
  cases1 = rbinom(H1, q1);
  cases2 = rbinom(H2, q2);
  cases3 = rbinom(H3, q3);
")

rproc <- Csnippet("
  int I = I1 + I2 + I3;
  int trans_S1[3], trans_S2[3], trans_S3[2], trans_I1[3], trans_I2[3], 
  trans_I3[2], trans_R1[3], trans_R2[3], trans_R3[2];

  double prob_S1[3], prob_I1[3], prob_R1[3], prob_S2[3], prob_I2[3], prob_R2[3],
  prob_S3[2], prob_I3[2], prob_R3[2];

  double xi = rgamma(sigma_xi, 1/sigma_xi);

  double kappa = (1 + beta11*cos(2*3.141593*t/52 + phi)) * xi;

  // Define rate
  prob_S1[0] = 1 - exp(-dt*beta1*kappa*I/N); // 0->1
  prob_S1[1] = 1 - exp(-delta1*dt);
  prob_S1[2] = exp(-delta1*dt) + exp(-dt*beta1*kappa*I/N) - 1;

  prob_I1[0] = 1 - exp(-gamma*dt);
  prob_I1[1] = 1 - exp(-delta1*dt);
  prob_I1[2] = exp(-gamma*dt)+exp(-delta1*dt) - 1;

  prob_R1[0] = 1 - exp(-omega*dt);  // E_1,t this goes back to S_1,(t+1)
  prob_R1[1] = 1 - exp(-delta1*dt);
  prob_R1[2] = exp(-omega*dt) + exp(-delta1*dt) - 1;

  prob_S2[0] = 1 - exp(-dt*beta2*kappa*I/N);
  prob_S2[1] = 1 - exp(-delta2*dt);
  prob_S2[2] = exp(-delta2*dt) + exp(-dt*beta2*kappa*I/N) - 1;

  prob_I2[0] = 1 - exp(-dt*gamma);
  prob_I2[1] = 1 - exp(-dt*delta2);
  prob_I2[2] = exp(-dt*gamma)+exp(-dt*delta2) - 1;

  prob_R2[0] = 1 - exp(-dt*omega);  // E_1,t this goes back to S_1,(t+1)
  prob_R2[1] = 1 - exp(-dt*delta2);
  prob_R2[2] = exp(-dt*omega) + exp(-dt*delta2) - 1;

  // For Age Group (3): Die first before transition;

  int S3mD, I3mD, R3mD;

  S3mD = rbinom(S3, 1-dt*mu); // S3 minus Death: mu is the death rate, so it's 1-mu here
  I3mD = rbinom(I3, 1-dt*mu);
  R3mD = rbinom(R3, 1-dt*mu);

  prob_S3[0] = 1-exp(-dt*beta3*kappa*I/N);
  prob_S3[1] = exp(-dt*beta3*kappa*I/N);

  prob_I3[0] = 1 - exp(-dt*gamma);
  prob_I3[1] = exp(-dt*gamma);

  prob_R3[0] = 1 - exp(-dt*omega);
  prob_R3[1] = exp(-dt*omega);

  // Transition
  // B: S->I
  // C: I->R
  // F: Aging: (1)->(2)->(3)
  // E: R->S
  // D: Death
  //// Note: Here S_1, S_2... are all old values from (t-1)
  rmultinom(S1, &prob_S1, 3, &trans_S1); // B, F, S-B-F
  rmultinom(I1, &prob_I1, 3, &trans_I1); // C, F, I-C-F
  rmultinom(R1, &prob_R1, 3, &trans_R1); // E, F, R-E-F

  rmultinom(S2, &prob_S2, 3, &trans_S2); // B, F, S-B-F
  rmultinom(I2, &prob_I2, 3, &trans_I2); // C, F, I-C-F
  rmultinom(R2, &prob_R2, 3, &trans_R2); // E, F, R-E-F

  rmultinom(S3mD, &prob_S3, 2, &trans_S3); // B, (S-D)-B
  rmultinom(I3mD, &prob_I3, 2, &trans_I3); // C, (I-D)-C
  rmultinom(R3mD, &prob_R3, 2, &trans_R3); // E, (R-D)-E

  S1 = trans_S1[2] + trans_R1[0] + rpois(4*1025.7); // Include Birth
  I1 = trans_I1[2] + trans_S1[0];
  R1 = trans_R1[2] + trans_I1[0];

  S2 = trans_S2[2] + trans_R2[0] + trans_S1[1]; // Include Aging
  I2 = trans_I2[2] + trans_S2[0] + trans_I1[1];
  R2 = trans_R2[2] + trans_I2[0] + trans_R1[1];

  S3 = trans_S3[1] + trans_R3[0] + trans_S2[1]; // Include Aging
  I3 = trans_I3[1] + trans_S3[0] + trans_I2[1];
  R3 = trans_R3[1] + trans_I3[0] + trans_R2[1];

  //Accumvar
  H1 += trans_S1[0];
  H2 += trans_S2[0];
  H3 += trans_S3[0];

  q1 = -1; 
  while(q1 < 0 || q1 > 1){
    q1 = rnorm(0.07, sigma_q);
  }

  q2 = -1; 
  while(q2 < 0 || q2 > 1){
    q2 = rnorm(0.07, sigma_q);
  }

  q3 = -1; 
  while(q3 < 0 || q3 > 1){
    q3 = rnorm(0.07, sigma_q);
  }
")

# full-estimate-rinit
rinit_full <- Csnippet("
  double m = N/(S10+I10+R10+S20+I20+R20+S30+I30+R30);
  I1=nearbyint(m*I10);
  I2=nearbyint(m*I20);
  I3=nearbyint(m*I30);
  S1=nearbyint(m*S10);
  S2=nearbyint(m*S20);
  S3=nearbyint(m*S30);
  R1=nearbyint(m*R10);
  R2=nearbyint(m*R20);
  R3=nearbyint(m*R30);
  H1 = 0;
  H2 = 0;
  H3 = 0;
")

# WWR's rinit
rinit_wwr <- Csnippet("
  S1=3876549;
  S2=57139612;
  S3=19573727;
  I1=30351;
  I2=871;
  I3=2550;
  R1=1315221;
  R2=302852;
  R3=131092;
  H1 = 0;
  H2 = 0;
  H3 = 0;
")

rinit_stocks <- Csnippet("
  S1=5.09484e+06;
  S2=5.73856e+07;
  S3=1.96976e+07;
  I1=2871;
  I2=639;
  I3=174;
  R1=124410;
  R2=57072;
  R3=9578;
  H1 = 0;
  H2 = 0;
  H3 = 0;
")

# x <- c(3876549,57139612,19573727,30351,871,2550,1315221,302852,131092)
# x/sum(x)

# define parameters (fixed)
params_fixed <- c(
  gamma = 1, 
  delta1 = 1/(5*52),
  delta2 = 1/(55*52), 
  alpha = 1/(78.86912*52), 
  mu = 0, 
  N = 82372825, 
  omega = 1/(1*52)
)
# Set to MLE
params_wwr <- c(
  params_fixed, 
  "beta1" = 11.48,
  "beta2" = 0.25,
  "beta3" = 0.35,
  "phi" = 0.14,
  "beta11" = 0.16,
  "sigma_q" = 0.021,
  "sigma_xi" = 66.89
)

params_full <- c(
  params_wwr,
  "S10" = 4.706102e-02,
  "S20" = 6.936707e-01,
  "S30" = 2.376236e-01,
  "I10" = 3.684589e-04,
  "I20" = 1.057388e-05,
  "I30" = 3.095681e-05,
  "R10" = 1.596669e-02,
  "R20" = 3.676601e-03,
  "R30" = 1.591447e-03
)

pt_full <- pomp::parameter_trans(
  log = c("beta1", "beta2", "beta3", "sigma_q", "sigma_xi"),
  logit = c("beta11"),
  barycentric = c("S10","I10","R10",
                  "S20","I20","R20",
                  "S30","I30","R30"),
  toEst = pomp::Csnippet("T_phi = logit(phi/(M_2PI));"),
  fromEst = pomp::Csnippet("phi = M_2PI*expit(T_phi);")
)

pt_wwr <- pomp::parameter_trans(
  log = c("beta1", "beta2", "beta3", "sigma_q", "sigma_xi"),
  logit = c("beta11"),
  toEst = pomp::Csnippet("T_phi = logit(phi/(M_2PI));"),
  fromEst = pomp::Csnippet("phi = M_2PI*expit(T_phi);")
)

dat <- read.table("real_rotavirus_metadata.txt") |>
  arrange(time)

statenames <- c(
  "S1", "I1", "R1", "H1",
  "S2", "I2", "R2", "H2",
  "S3", "I3", "R3", "H3",
  "q1", "q2", "q3"
)

sir_wwr <- pomp(
  data = dat,
  times = "time",
  t0 = 0,
  dmeasure = dmeas,
  rmeasure = rmeas,
  rprocess = discrete_time(step.fun = rproc, delta.t = 1/4),
  statenames = statenames,
  paramnames = names(params_wwr),
  accumvars = c("H1", "H2", "H3"),
  rinit = rinit_wwr,
  partrans = pt_wwr,
  params = params_wwr
)

sir_full <- pomp(
  sir_wwr,
  rinit = rinit_full,
  params = params_full,
  partrans = pt_full,
  statenames = statenames,
  paramnames = names(params_full)
)

sir_sbh_init <- pomp(
  sir_wwr,
  t0 = 1-6*52,
  rinit = rinit_stocks,
  statenames = statenames
) 
@

@@

<<maximization1, echo=F, warning=F, message=F>>=
set.seed(123)
sir_box <- list(
  beta1 = c(10, 15),
  beta2 = c(0.2, 0.4),
  beta3 = c(0.3, 0.5),
  phi = c(0.01, 0.3),
  beta11 = c(0.1, 0.2),
  sigma_q = c(0.001, 0.1),
  sigma_xi = c(65, 70),
  S10 = c(0.01, 0.05),
  I10 = c(0.0001, 0.0005),
  R10 = c(0.01, 0.05),
  S20 = c(0.01, 0.7),
  I20 = c(0.000001, 0.00002),
  R20 = c(0.001, 0.005),
  S30 = c(0.01, 0.5),
  I30 = c(0.000001, 0.0001),
  R30 = c(0.0001, 0.005),
  gamma = c(1, 1),
  delta1 = c(1/(5*52), 1/(5*52)),
  delta2 = c(1/(55*52), 1/(55*52)),
  alpha = c(1/(78.86912*52), 1/(78.86912*52)),
  mu = c(0, 0),
  N = c(82372825, 82372825),
  omega = c(1/(1*52), 1/(1*52))
)

design_matrix <- runif_design(
  lower = sapply(sir_box, `[`, 1),
  upper = sapply(sir_box, `[`, 2),
  n = ncores
)

bake(file = paste0(out_dir, "mif2_maximization/full/round_01/mifs_global_01.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      sir_full,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(
        beta1 = 0.01, beta2 = 0.01, beta3 = 0.01, beta11 = 0.01, 
        phi = 0.01, sigma_q = 0.01, sigma_xi = 0.01,
        S10 = ivp(0.24), I10 = ivp(0.24), R10 = ivp(0.24),
        S20 = ivp(0.24), I20 = ivp(0.24), R20 = ivp(0.24),
        S30 = ivp(0.24), I30 = ivp(0.24), R30 = ivp(0.24)
      ),
      cooling.type = "geometric",
      Nmif = NMIF,
      params = design_matrix[i, ]
    ) 
  }
}) -> mifs_global_01

bake(file = paste0(out_dir, "mif2_maximization/full/round_01/el_01.rds"),{
  eval_logLik_pomp(mifs_global_01, ncores, NP_MIF, 4248930)
}) ->  el_01

@

@@

<<maximization2, echo=F, warning=F, message=F>>=
set.seed(1234)
sd_02 = 2/3
starting_values_02 <- choose_n_fits(el_01, top_n_fits, ncores)

bake(file = paste0(out_dir,"mif2_maximization/full/round_02/mifs_global_02.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      mifs_global_01[[1]],
      rw.sd = rw_sd(
        beta1 = 0.01*sd_02, beta2 = 0.01*sd_02, beta3 = 0.01*sd_02,
        beta11 = 0.01*sd_02, phi = 0.01*sd_02,
        sigma_q = 0.01*sd_02, sigma_xi = 0.01*sd_02,
        S10 = ivp(0.24)*sd_02, I10 = ivp(0.24)*sd_02, R10 = ivp(0.24)*sd_02,
        S20 = ivp(0.24)*sd_02, I20 = ivp(0.24)*sd_02, R20 = ivp(0.24)*sd_02,
        S30 = ivp(0.24)*sd_02, I30 = ivp(0.24)*sd_02, R30 = ivp(0.24)*sd_02
      ),
      params = starting_values_02[[i]]
    ) 
  }
}) ->  mifs_global_02

bake(file = paste0(out_dir, "mif2_maximization/full/round_02/el_02.rds"),{
  eval_logLik_pomp(mifs_global_02, ncores, NP_MIF, seed=4248930)
}) ->  el_02
@

@@

<<maximization3, echo=F, warning=F, message=F>>=
set.seed(12345)
sd_03 = (2/3)^2
starting_values_03 <- choose_n_fits(el_02, top_n_fits, ncores)

bake(file = paste0(out_dir, "mif2_maximization/full/round_03/mifs_global_03.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      mifs_global_02[[1]],
      rw.sd = rw_sd(
        beta1 = 0.01*sd_03, beta2 = 0.01*sd_03, beta3 = 0.01*sd_03,
        beta11 = 0.01*sd_03, phi = 0.01*sd_03,
        sigma_q = 0.01*sd_03, sigma_xi = 0.01*sd_03,
        S10 = ivp(0.24)*sd_03, I10 = ivp(0.24)*sd_03, R10 = ivp(0.24)*sd_03,
        S20 = ivp(0.24)*sd_03, I20 = ivp(0.24)*sd_03, R20 = ivp(0.24)*sd_03,
        S30 = ivp(0.24)*sd_03, I30 = ivp(0.24)*sd_03, R30 = ivp(0.24)*sd_03
      ),
      params = starting_values_03[[i]]
    ) 
  }
}) -> mifs_global_03

bake(file = paste0(out_dir,"mif2_maximization/full/round_03/el_03.rds"),{
  eval_logLik_pomp(mifs_global_03, ncores, NP_MIF, seed = 4248930)
}) ->  el_03
@

<<estimated-mle, echo=FALSE, message=FALSE, warning=FALSE>>=
# Can be changed if more rounds needed
est_mle_vec <- el_03[which.max(el_03$logLik),]
@

@@

<<maximization1-Stocks_rinit, echo=F, warning=F, message=F>>=
set.seed(123)
sir_box_sbh <- list(
  beta1 = c(10, 15),
  beta2 = c(0.2, 0.4),
  beta3 = c(0.3, 0.5),
  phi = c(0.01, 0.3),
  beta11 = c(0.1, 0.2),
  sigma_q = c(0.001, 0.1),
  sigma_xi = c(65, 70),
  gamma = c(1, 1),
  delta1 = c(1/(5*52), 1/(5*52)),
  delta2 = c(1/(55*52), 1/(55*52)),
  alpha = c(1/(78.86912*52), 1/(78.86912*52)),
  mu = c(0, 0),
  N = c(82372825, 82372825),
  omega = c(1/(1*52), 1/(1*52))
)

design_matrix_sbh <- runif_design(
  lower = sapply(sir_box_sbh, `[`, 1),
  upper = sapply(sir_box_sbh, `[`, 2),
  n = ncores
)

bake(file = paste0(out_dir, "mif2_maximization/SBH_init/round_01/mifs_global_sbh_01.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      sir_sbh_init,
      Np = NP_MIF,
      cooling.fraction.50 = 0.5,
      rw.sd = rw_sd(
        beta1 = 0.01, beta2 = 0.01, beta3 = 0.01, beta11 = 0.01, 
        phi = 0.01, sigma_q = 0.01, sigma_xi = 0.01
      ),
      cooling.type = "geometric",
      Nmif = NMIF,
      params = design_matrix_sbh[i, ]
    ) 
  }
}) -> mifs_global_sbh_01

bake(file = paste0(out_dir, "mif2_maximization/SBH_init/round_01/sbh_el_01.rds"),{
  eval_logLik_pomp(mifs_global_sbh_01, ncores, NP_MIF, 4248930)
}) -> sbh_el_01

@

@@

<<maximization2-Stocks_rinit, echo=F, warning=F, message=F>>=
set.seed(1234)
sd_02 = 2/3
starting_values_sbh_02 <- choose_n_fits(sbh_el_01, top_n_fits, ncores)

bake(file = paste0(out_dir,"mif2_maximization/SBH_init/round_02/mifs_global_sbh_02.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      mifs_global_sbh_01[[1]],
      rw.sd = rw_sd(
        beta1 = 0.01*sd_02, beta2 = 0.01*sd_02, beta3 = 0.01*sd_02,
        beta11 = 0.01*sd_02, phi = 0.01*sd_02,
        sigma_q = 0.01*sd_02, sigma_xi = 0.01*sd_02
      ),
      params = starting_values_sbh_02[[i]]
    ) 
  }
}) ->  mifs_global_sbh_02

bake(file = paste0(out_dir, "mif2_maximization/SBH_init/round_02/sbh_el_02.rds"),{
  eval_logLik_pomp(mifs_global_sbh_02, ncores, NP_MIF, seed=4248930)
}) -> sbh_el_02
@

@@

<<maximization3-Stocks_rinit, echo=F, warning=F, message=F>>=
set.seed(12345)
sd_03 = (2/3)^2
starting_values_sbh_03 <- choose_n_fits(sbh_el_02, top_n_fits, ncores)

bake(file = paste0(out_dir, "mif2_maximization/SBH_init/round_03/mifs_global_sbh_03.rds"),{ 
  plan(multicore, workers = ncores)
  foreach(i = 1:ncores, .options.future = list(seed = 652643293)) %dofuture% {
    pomp::mif2(
      mifs_global_sbh_02[[1]],
      rw.sd = rw_sd(
        beta1 = 0.01*sd_03, beta2 = 0.01*sd_03, beta3 = 0.01*sd_03,
        beta11 = 0.01*sd_03, phi = 0.01*sd_03,
        sigma_q = 0.01*sd_03, sigma_xi = 0.01*sd_03
      ),
      params = starting_values_sbh_03[[i]]
    ) 
  }
}) -> mifs_global_sbh_03

bake(file = paste0(out_dir,"mif2_maximization/SBH_init/round_03/sbh_el_03.rds"),{
  eval_logLik_pomp(mifs_global_sbh_03, ncores, NP_MIF, seed = 4248930)
}) ->  sbh_el_03
@

<<estimated-mle_stocks_init, echo=FALSE, message=FALSE, warning=FALSE>>=
# Can be changed if more rounds needed
est_mle_vec_sbh <- sbh_el_03[which.max(sbh_el_03$logLik),]
@

@@

<<sim-100, echo=F, warning=F, message=F>>=
bake(file = paste0(out_dir, "pfilter/simulated_dataset/ovov_sim_100.rds"),{
  ovov_sim_computation <- list()
  i <- 1
  repeat {
    # Simulation
    sim_data <- simulate(sir_wwr, format = "data.frame") |>
      subset(select=c("time", "cases1", "cases2", "cases3"))

    if (!any(sim_data == 0)) {
      ovov_sim_computation[[length(ovov_sim_computation) + 1]] <- sim_data
      i <- i + 1  
    } else {
      print(paste("Condition not met for iteration", i, "- retrying"))
    }

    if (length(ovov_sim_computation) >= sim_rep) break  
  }
  ovov_sim_computation
}) -> ovov_sim_100
@

<<pfilter_sim_100, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
bake(file = paste0(out_dir, "pfilter/simulated_dataset/pfilter_ovov_sim_100.rds"),{
  pfilter_obj <- list()
  pfilter_ovov_sim_computation <- c()

  for(i in 1:length(ovov_sim_100)){
    pomp(
      data = ovov_sim_100[[i]],
      times = "time",
      t0 = 0,
      dmeasure = dmeas,
      rmeasure = rmeas,
      rprocess = discrete_time(step.fun = rproc, delta.t = 1/4),
      statenames = statenames,
      paramnames = names(params_wwr),
      accumvars = c("H1", "H2", "H3"),
      rinit = rinit_wwr,
      partrans = pt_wwr,
      params = params_wwr
    )  -> pfilter_obj[[i]]

    plan(multicore, workers = ncores)

    foreach(j = 1:ncores, .combine = c,
      .options.future = list(seed = 998468235L)
    ) %dofuture% {
      pfilter(
        pfilter_obj[[i]], Np = NP_MIF, save.states = FALSE, params = params_wwr
      )
    } -> pfs
    pfilter_ovov_sim_computation[i] <- logmeanexp(logLik(pfs))  
  }
  pfilter_ovov_sim_computation
}) ->  pfilter_ovov_sim_100
@

<<palfilter_sim_100, echo=FALSE, message=FALSE, warning=FALSE, results='hide'>>=
prop <- 420
init_dist <- c(
  3876549, 30351, 1315221, 57139612, 871, 302852, 19573727, 2550, 131092
)
initial_guess <- c(11.48, 0.25, 0.35, 0.14, 0.16, 0.021, 66.89)

bake(file = paste0(out_dir, "pfilter/simulated_dataset/palfilter_ovov_sim_100.rds"),{
  plan(multicore, workers = ncores)
  foreach(
    i = 1:length(ovov_sim_100), .combine = c, .options.future = list(seed = 998468235L)
  ) %dofuture% {
    pal_sim_dat <- ovov_sim_100[[i]][,-1] |> as.matrix()
    lik_list <- rotavirus_SMC_qropxi(
      init_dist = init_dist, 
      y_obs = pal_sim_dat, 
      m = 9, 
      regular_params = c(
        initial_guess[1], initial_guess[2], initial_guess[3], initial_guess[4],
        initial_guess[5]
      ), 
      gamma_par = c(initial_guess[7], 1/initial_guess[7]), 
      norm_par = c(0.07, initial_guess[6]), 
      prop = t(prop), 
      n_particles = NP_MIF,
      ncores = ncores
    )
    lik_list$log_lik
  }
}) -> palfilter_ovov_sim_100

# lik_list$ll_storage |> sum()
@

<<pfilter_wwr, echo=FALSE, message=FALSE, warning=FALSE>>=
bake(file = paste0(out_dir, "pfilter/original_dataset/pfilter_wwr.rds"),{
  plan(multicore, workers = ncores)
  foreach(
    i = 1:ncores, .combine = c, .options.future = list(seed = 998468235L)
  ) %dofuture% {
      pfilter(sir_wwr, Np = NP_MIF, params = params_wwr)
    } 
}) ->  pfilter_wwr

bake(file = paste0(out_dir, "pfilter/original_dataset/pfilter_wwr_cond_lik.rds"),{
  pfilter_wwr_cond_lik_computation <- list()
  for(k in 1:ncores){
    pfilter_wwr_cond_lik_computation[[k]] <- pfilter_wwr[[k]]@cond.logLik
  }
  pfilter_wwr_cond_lik_computation
}) -> pfilter_wwr_cond_lik
@

<<palfilter_wwr, echo=FALSE>>=
realdat <- as.matrix(dat[,-1])

bake(file = paste0(out_dir,"pfilter/original_dataset/palfilter_lik_list_wwr.rds"),{
 set.seed(1223)
 rotavirus_SMC_qropxi(
    init_dist = init_dist, 
    y_obs = realdat, 
    m = 9, 
    regular_params = c(
      initial_guess[1], initial_guess[2], initial_guess[3], initial_guess[4],
      initial_guess[5]
    ), 
    gamma_par = c(initial_guess[7], 1/initial_guess[7]), 
    norm_par = c(0.07, initial_guess[6]), 
    prop = t(prop), 
    n_particles = NP_MIF, 
    ncores = ncores
  ) 
}) -> palfilter_lik_list_wwr
@

<<PAL-MLE-from-mif2, echo=FALSE, message=FALSE, warning=FALSE>>=
prop <- 420
realdat <- as.matrix(dat[,-1])
init_dist_mif2_mle <- c(
  pop*as.numeric(est_mle_vec["S10"]), 
  pop*as.numeric(est_mle_vec["I10"]), 
  pop*as.numeric(est_mle_vec["R10"]),
  pop*as.numeric(est_mle_vec["S20"]),
  pop*as.numeric(est_mle_vec["R20"]),
  pop*as.numeric(est_mle_vec["I20"]),
  pop*as.numeric(est_mle_vec["S30"]),
  pop*as.numeric(est_mle_vec["R30"]),
  pop*as.numeric(est_mle_vec["I30"])
)

initial_guess_mif2_mle <- c(
  as.numeric(est_mle_vec["beta1"]), 
  as.numeric(est_mle_vec["beta2"]), 
  as.numeric(est_mle_vec["beta3"]),
  as.numeric(est_mle_vec["phi"]),
  as.numeric(est_mle_vec["beta11"]),
  as.numeric(est_mle_vec["sigma_q"]),
  as.numeric(est_mle_vec["sigma_xi"])
)

bake(file = paste0(out_dir,"pfilter/original_dataset/palfilter_mif2_mle.rds"),{
 set.seed(1223)
 rotavirus_SMC_qropxi(
    init_dist = init_dist_mif2_mle, 
    y_obs = realdat, 
    m = 9, 
    regular_params = c(
      initial_guess_mif2_mle[1], 
      initial_guess_mif2_mle[2], 
      initial_guess_mif2_mle[3], 
      initial_guess_mif2_mle[4],
      initial_guess_mif2_mle[5]
    ), 
    gamma_par = c(initial_guess_mif2_mle[7], 1/initial_guess_mif2_mle[7]), 
    norm_par = c(0.07, initial_guess_mif2_mle[6]), 
    prop = t(prop), 
    n_particles = NP_MIF, 
    ncores = ncores
  ) 
}) -> palfilter_mif2_mle
@

>>>>

\documentclass[10pt]{article}\usepackage[]{graphicx}\usepackage[]{xcolor}
% maxwidth is the original width if it is less than linewidth
% otherwise use linewidth (to make sure the graphics do not exceed the margin)
\makeatletter
\def\maxwidth{ %
  \ifdim\Gin@nat@width>\linewidth
    \linewidth
  \else
    \Gin@nat@width
  \fi
}
\makeatother

\definecolor{fgcolor}{rgb}{0.345, 0.345, 0.345}
\newcommand{\hlnum}[1]{\textcolor[rgb]{0.686,0.059,0.569}{#1}}%
\newcommand{\hlsng}[1]{\textcolor[rgb]{0.192,0.494,0.8}{#1}}%
\newcommand{\hlcom}[1]{\textcolor[rgb]{0.678,0.584,0.686}{\textit{#1}}}%
\newcommand{\hlopt}[1]{\textcolor[rgb]{0,0,0}{#1}}%
\newcommand{\hldef}[1]{\textcolor[rgb]{0.345,0.345,0.345}{#1}}%
\newcommand{\hlkwa}[1]{\textcolor[rgb]{0.161,0.373,0.58}{\textbf{#1}}}%
\newcommand{\hlkwb}[1]{\textcolor[rgb]{0.69,0.353,0.396}{#1}}%
\newcommand{\hlkwc}[1]{\textcolor[rgb]{0.333,0.667,0.333}{#1}}%
\newcommand{\hlkwd}[1]{\textcolor[rgb]{0.737,0.353,0.396}{\textbf{#1}}}%
\let\hlipl\hlkwb

\usepackage{framed}
\makeatletter
\newenvironment{kframe}{%
 \def\at@end@of@kframe{}%
 \ifinner\ifhmode%
  \def\at@end@of@kframe{\end{minipage}}%
  \begin{minipage}{\columnwidth}%
 \fi\fi%
 \def\FrameCommand##1{\hskip\@totalleftmargin \hskip-\fboxsep
 \colorbox{shadecolor}{##1}\hskip-\fboxsep
     % There is no \\@totalrightmargin, so:
     \hskip-\linewidth \hskip-\@totalleftmargin \hskip\columnwidth}%
 \MakeFramed {\advance\hsize-\width
   \@totalleftmargin\z@ \linewidth\hsize
   \@setminipage}}%
 {\par\unskip\endMakeFramed%
 \at@end@of@kframe}
\makeatother

\definecolor{shadecolor}{rgb}{.97, .97, .97}
\definecolor{messagecolor}{rgb}{0, 0, 0}
\definecolor{warningcolor}{rgb}{1, 0, 1}
\definecolor{errorcolor}{rgb}{1, 0, 0}
\newenvironment{knitrout}{}{} % an empty environment to be redefined in TeX

\usepackage{alltt}
\usepackage{color}
\usepackage{graphicx} % Required for inserting
\usepackage{booktabs} % For better-quality horizontal lines
\usepackage{multirow}
\usepackage{makecell}
% \usepackage{threeparttable} % For table notes
\usepackage{array} % For table formatting
\usepackage{subcaption}
\usepackage{setspace} % For double-line spacing
\usepackage{float} 
\usepackage{tabularx}
\usepackage[round]{natbib}
\usepackage[font={small,it}]{caption}
\usepackage[utf8]{inputenc}
\usepackage{mathtools}
\usepackage{fullpage}
\usepackage{verbatim}
\usepackage{amsfonts,amssymb,amsthm}
\usepackage{url}
\usepackage[colorlinks=true,citecolor=blue]{hyperref}
% Using \doublespacing in the preamble 
% changes the text to double-line spacing
% \doublespacing

@@

\bibliographystyle{apalike}
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\begin{document}

@@

@@

@@

@@

@@

@@

@@

@@

@@

\begin{table}[ht] % The placement specifier can be [h!tbp]
\centering % Centers the table
\caption{AIC for the OvOv rotavirus model, computed using two filtering methods. PAL is the Poisson approximate likelihood, implemented using the code of WWR. PF is the particle filter, implemented using the code of \citet{pomppackagepaper}. The first two lines are from WWR, Table~1, and lines 3--5 are our own computations. We used 50000 particle for both PF and PAL. PF was repeated 36 times to reduce the Monte Carlo variance, but this step was not necessary for PAL.
\ed{SHOULD WE HAVE A LINE FOR PAL AT THE IF2 MLE FROM LINE 5? PERHAPS WE CAN'T READILY MAXIMIZE PAL FOR THIS MODEL (OR CAN WE?) BUT SOME SORT OF COMPARISON WOULD BE HELPFUL}
} 
\label{tab:ovovrealdata} % For referencing this table elsewhere in your text
\begin{tabular}{lllc} 
  \hline
  & Method & Result & AIC  \\
  \hline
  1. & PAL & Table~1 of WWR & 13778.08
  \\
  2. & PF & Table~1 of WWR, originally from SBH &  20134.38
  \\
  3. & PAL & Model, data and MLE from WWR &
    13814.57 
  \\
  4. & PF & Model, data and MLE from WWR &
    25785.52
  \\
  5. & PF & Data from WWR, estimated initial values &
     143875.53
  \\
  6. & PAL & Data from WWR, estimated initial values &
     15376.67
  \\
 \hline
\end{tabular}
\end{table}
